---
title: "R codes for statistical analyses of MPRNA-IP data"
author: "Taeyoung Hwang, PhD"
date: "February 7, 2024"
output: github_document
---

```{r, message=F}
library(cowplot)
library(ggrepel)
library(ggpubr)

library(dplyr)
library(reshape2)

library(edgeR)
```

```{r}
theme_set(theme_cowplot())
source("nar_functions.R")
```

# 1. PUM2-RNA MPRNA-IP

### Load raw count data and tile annotations.
```{r}
load("RawData/git_PUM2_raw.Rdata")
BARCODE_NUM <- 22
REPLICATE_NUM <- 3
```

### Filtering and pooling
```{r}
pooled.count <- oligoFilter(oligo.count, inputFilterBy="rpm", inputTh=1, ipFilterBy="rpm", ipTh=1) %>% countPooler()
```

```{r}
pooled.fastqc <- fastqc[match(colnames(pooled.count), fastqc$SampleID),]
pooled.fastqc <- droplevels(pooled.fastqc) 
```

### Enrichment

Limma + Voom
```{r}
pooled.model <- model.matrix(data= pooled.fastqc, ~ 0 + Pulldown + Repl)
edgeR.obj <- DGEList(pooled.count)
edgeR.obj <- calcNormFactors(edgeR.obj, method="TMM")
y <- voom(edgeR.obj, pooled.model, plot = T)
fit <- lmFit(y, pooled.model)
contr <- makeContrasts(PulldownIP - PulldownInput, levels = colnames(coef(fit)))
out <- contrasts.fit(fit, contr)
out <- eBayes(out)
out.table <- topTable(out, sort.by = "P", n = Inf)
table(out.table$adj.P.Val<0.01, out.table$logFC>0)
```

Significant tiles
```{r}
tile.table <- data.frame(tileID=tile.annot$tileID, stringsAsFactors = F)
idx <- match(tile.table$tileID, rownames(out.table))
tile.table$rpm <- 2^out.table$AveExpr[idx]
tile.table$log2FC <- out.table$logFC[idx]
tile.table$stat <- out.table$t[idx]
tile.table$pval <- out.table$P.Value[idx]
tile.table$padj <- out.table$adj.P.Val[idx]
tile.table$sig <- factor(tile.table$padj<0.01, levels=c("TRUE", "FALSE"), labels=c("Sig", "Insig"))

tile.table$enriched <- NA
tile.table$enriched[!is.na(tile.table$stat)] <- "Insig"
tile.table$enriched[which(tile.table$sig=="Sig" & tile.table$log2FC>0)] <- "Enriched"
tile.table$enriched[which(tile.table$sig=="Sig" & tile.table$log2FC<0)] <- "Depleted"
table(tile.table$enriched)
```

MA plot
```{r}
p <- ggplot(subset(tile.table, !is.na(log2FC)), aes(x=log10(rpm), y=log2FC))
p <- p + geom_point(data=subset(tile.table, enriched=="Insig"), col="gray", alpha=0.5)
p <- p + geom_point(data=subset(tile.table, enriched=="Depleted"), col="blue", alpha=0.8)
p <- p + geom_point(data=subset(tile.table, enriched=="Enriched"), col="red", alpha=0.8)
p <- p + geom_hline(yintercept = 0, linetype="dashed")
p <- p + xlab("Log10 (Mean RPM)") + ylab("Log2 (IP/Input)")
p
```

### Analysis
PRE effect on fold change
```{r}
temp <- subset(tile.table, !is.na(log2FC))
temp$pre.num <- tile.annot$pre.num[match(temp$tileID, tile.annot$tileID)]
table(temp$pre.num)

p <- ggviolin(temp, x="pre.num", y="log2FC", color="pre.num", palette=c("black", "blue", "red"), alpha=1, add="jitter", add.params=list(size=1, alpha=0.7))
p <- p + stat_compare_means(comparisons = list( c("0", "1"), c("1", "2") ), size=5)
p <- p + geom_hline(yintercept = 0, linetype="dashed")
p
```

GC %
```{r}
temp <- cbind(tile.annot, tile.table[,-1])
temp <- subset(temp, !is.na(enriched))
p <- ggplot(temp, aes(col=enriched, gc))
p <- p + stat_ecdf()
p
```

# 2. MS2-MCP MPRNA-IP

### Load raw count data and tile annotations.
```{r}
load("RawData/git_ms2_raw.Rdata")
BARCODE_NUM <- 50
REPLICATE_NUM <- 3
```

Select Input and HA pulldown
```{r}
oligo.count <- subset(oligo.count, sample %in% c("rep1in", "rep1ha","rep2in", "rep2ha","rep3in", "rep3ha"))
oligo.count$pulldown <- gsub("HA", "IP" , oligo.count$pulldown)
```

Tidy
```{r}
oligo.count <- oligo.count[, c("tileID", "barcodeID", "repl", "pulldown", "raw", "rpm")]
rownames(oligo.count) <- 1:nrow(oligo.count)
```

### Normalization
```{r}
temp <- dcast(oligo.count, tileID+barcodeID~repl+pulldown, value.var="raw")
count.normed <- edgeR::calcNormFactors(edgeR::DGEList(as.matrix(temp[, -c(1,2)])), method="TMM")
count.normed <- edgeR::cpm(count.normed)

temp <- melt(cbind(temp[, c(1,2)], count.normed), id.vars=c("tileID", "barcodeID"), value.name = "normed", variable.name = "sampleID")
idx <- match(paste(oligo.count$tileID, oligo.count$barcodeID, oligo.count$repl, oligo.count$pulldown, sep="_") , paste(temp$tileID, temp$barcodeID, temp$sampleID, sep="_"))
oligo.count$normed <- temp$normed[idx]
```

### Single mutation

Single mutation table
```{r}
single.table <- subset(tile.annot, group=="Single")[, c("tileID", "mutType", "mutLoc")]
temp <- strsplit(single.table$mutType, split="to")
single.table$seqWT <- sapply(temp, "[[", 1)
single.table$seqMT <- sapply(temp, "[[", 2)
```

Limma for individual replicates
```{r}
temp <- data.frame(mutID=paste(single.table$mutLoc, single.table$mutType, sep="_"), tileID.mut=single.table$tileID, tileID.wt=subset(tile.annot, group=="WT")$tileID)

single.rep1 <- runLimma.singleMut(temp, "Rep1", 1)
single.rep2 <- runLimma.singleMut(temp, "Rep2", 1)
single.rep3 <- runLimma.singleMut(temp, "Rep3", 1)
```

Merge and p-value based on rank
```{r}
temp <- mergeLimmaResults(list(single.rep1$limma, single.rep2$limma, single.rep3$limma))
single.table <- cbind(single.table, temp)
rownames(single.table) <- rownames(temp)
```

Result
```{r}
temp <- single.table
temp$seqWT <- gsub("T", "U", temp$seqWT)
temp$seqMT <- gsub("T", "U", temp$seqMT)
temp$xlab <- paste(temp$mutLoc, temp$seqWT, sep=":")
temp$xlab <- factor(temp$xlab, levels=unique(temp$xlab))
temp$Sequence <- temp$seqMT
temp$FDR <- cut(temp$qval, breaks = c(0,0.1,1), labels=c("<=0.1", "<=1"))

p <- ggplot(temp, aes(x=xlab, y=2^mean.logFC, col=FDR, shape=Sequence))
p <- p + geom_segment(linewidth=0.5, aes(xend=xlab), yend=0, color='grey80')
p <- p + geom_point(size=3)
p <- p + scale_color_manual(values = c("red", "black"))
p <- p + geom_hline(yintercept = 1, linetype="dashed")
p <- p + theme(axis.text.x=element_text(size=15, angle=70, hjust=1, vjust=1, color = "black"))
p <- p + xlab("Location on MS2") + ylab(expression(Delta* " (HA / Input)"))
p
```


### Stem mutation effect
Stem mutant table
```{r}
# stem mutant table was pre-made.
```

Limma for individual replicates
```{r}
stem.rep1 <- runLimma.stemMut(stem.table, "Rep1", 1)
stem.rep2 <- runLimma.stemMut(stem.table, "Rep2", 1)
stem.rep3 <- runLimma.stemMut(stem.table, "Rep3", 1)
```

Merge and p-value based on rank of t-statistics
```{r}
temp <- mergeLimmaResults(list(stem.rep1$limma, stem.rep2$limma, stem.rep3$limma))
stem.table <- cbind(stem.table, temp)
rownames(stem.table) <- rownames(temp)
```

Results
```{r}
temp <- stem.table
temp$stemEffect <- 2^(temp$mean.logFC)
temp$FDR <- cut(temp$qval, breaks = c(0,0.1,1))

p <- ggpubr::ggdotchart(temp, x = "stemID", y = "stemEffect", col="FDR", dot.size=5, add = "segments", palette=c("red", "black"))
p <- p + geom_hline(yintercept = c(1,2), linetype="dashed")
p
```

Examples
```{r}
temp <- stem_log2FC_tabler("stem:8")
table(temp$tileID) # WT: 142, 5' mut: 76, 3' mut: 107, Comp. mut: 146
p <- ggpubr::ggboxplot(temp, x = "tileID", y = "log2FC", palette = "jco", add="jitter", add.params=list(size=2), outlier.shape=NA, legend = "none")
p <- p + ggpubr::stat_compare_means(comparisons = list( c("WT", "5' mut"), c("WT", "3' mut"), c("WT", "Comp. mut") ))
p <- p + geom_hline(yintercept = mean(subset(temp, tileID=="WT")$log2FC), linetype="dashed")
p # p-value: 1.4e-5, 1.1e-14, 0.047
```

# 3. hTR-FlagTERT MPRNA-IP

### Load raw count data and tile annotations.
```{r}
load("RawData/git_hTR_raw.Rdata")

BARCODE_NUM <- 25
REPLICATE_NUM <- 3
```

### Tile Enrichment

```{r}
temp <- cbind(oligo.count, tile.annot[match(oligo.count$tileID, tile.annot$tileID),-1]) 
temp <- subset(temp, type %in% c("Random", "Control"))

temp <- dcast(temp, type+coord+tileID+barcodeID+repl~pulldown, value.var = "rpm")
temp$log2fc <- log2((temp$IP+1) / (temp$Input+1))
temp <- subset(temp, Input>=1 & IP>=1) # RPM>=1
table(temp$coord) # 32, 39, 60, 5679
temp$fc <- 2^(temp$log2fc)
wilcox.test(subset(temp, coord=="Random")$fc, subset(temp, coord=="hTR:34-190")$fc) # p-value = 6.332e-06
wilcox.test(subset(temp, coord=="Random")$fc, subset(temp, coord=="hTR:210-366")$fc) # p-value < 2.2e-16
wilcox.test(subset(temp, coord=="Random")$fc, subset(temp, coord=="hTR:295-451")$fc) # p-value = 0.2652
wilcox.test(subset(temp, coord=="hTR:34-190")$fc, subset(temp, coord=="hTR:210-366")$fc) # p-value = 0.0007099

temp1 <- dcast(temp, coord~. , value.var="fc", mean)
temp1$se <- dcast(temp, coord~. , value.var="fc", function(x) {sd(x)/sqrt(length(x))})[,"."]
colnames(temp1) <- c("Tile", "Mean", "SE")
temp1$Tile <- factor(temp1$Tile, levels=c("Random", "hTR:34-190", "hTR:210-366", "hTR:295-451"), labels = c("Random", "Tile 1", "Tile 2", "Tile 3"))

p <- ggplot(temp1, aes(x=Tile, y=Mean, fill=Tile)) 
p <- p + geom_bar(stat="Identity")
p <- p + geom_errorbar(aes(ymin=Mean-SE, ymax=Mean+SE), linewidth=0.8, width=0.5)
p <- p + scale_fill_manual(values=alpha(c("gray50", "orange", "red", "gray10"), 0.8))
p <- p + geom_hline(yintercept = 1, linetype="dashed")
p

# effect size
temp1 # tile1: 1.65, tile2: 2.25
```

### Normalization
```{r}
temp <- dcast(oligo.count, tileID+barcodeID~repl+pulldown, value.var="raw")
count.normed <- edgeR::calcNormFactors(edgeR::DGEList(as.matrix(temp[, -c(1,2)])), method="TMM")
count.normed <- edgeR::cpm(count.normed)

temp <- melt(cbind(temp[, c(1,2)], count.normed), id.vars=c("tileID", "barcodeID"), value.name = "normed", variable.name = "sampleID")
idx <- match(paste(oligo.count$tileID, oligo.count$barcodeID, oligo.count$repl, oligo.count$pulldown, sep="_") , paste(temp$tileID, temp$barcodeID, temp$sampleID, sep="_"))
oligo.count$normed <- temp$normed[idx]
```

### Deletion and replace

Deletion (and replace) table
```{r}
deletion.table <- subset(tile.annot, coord %in% c("hTR:34-190", "hTR:210-366") & mutation %in% c("Deletion", "Replace"))

deletion.table$mutID <- deletion.table$tileID
deletion.table$tileID.mut  <- deletion.table$tileID
temp <- subset(tile.annot, type=="Control")
deletion.table$tileID.wt <- temp$tileID[match(deletion.table$coord, temp$coord)]
```

Limma for individual replicates
```{r}
deletion.rep1 <- runLimma.singleMut(deletion.table, "Rep1", 1)
deletion.rep2 <- runLimma.singleMut(deletion.table, "Rep2", 1)
deletion.rep3 <- runLimma.singleMut(deletion.table, "Rep3", 1)
```

Merge and p-value based on rank
```{r}
temp <- mergeLimmaResults(list(deletion.rep1$limma, deletion.rep2$limma, deletion.rep3$limma))
deletion.table <- cbind(deletion.table, temp)
rownames(deletion.table) <- rownames(temp)
```

Results
```{r}
temp <- deletion.table
temp$region <- factor(temp$region, levels=sort(unique(temp$region)))
temp$Tile <- factor(temp$coord, levels=c("hTR:34-190", "hTR:210-366"), labels=c("Pesudoknot", "CR4/CR5"))
temp$Significance <- cut(temp$qval, breaks=c(0,0.1,0.3,1))
table(temp$Significance) # 2 significant

temp1 <- subset(temp, mutation=="Deletion")
temp2 <- subset(temp, mutation=="Replace")
all(temp1$bpLoc5prime == temp1$bpLoc5prime) & all(temp1$bpLoc3prime == temp1$bpLoc3prime)

temp1$logFC2 <- temp2$mean.logFC
temp1$Significance2 <- temp2$Significance

p <- ggplot(temp1, aes(x=region, y=mean.logFC, col=Significance), shape="circle")
p <- p + geom_point(size=3)
p <- p + geom_point(aes(y=logFC2, col=Significance2), size=3, shape="triangle")
#p <- p + scale_color_manual(values=c("red", "orange", "black")) # no points for orange
p <- p + scale_color_manual(values=c("red", "black")) 
p <- p + geom_segment(size=0.7, aes(xend=region, yend=logFC2), color='blue')
p <- p + geom_hline(yintercept = 0, linetype="dashed")
p <- p + background_grid() 
p
```

### Stem mutation

Stem mutation table
```{r}
temp <- subset(tile.annot, type %in% c("Mut1bp", "Mut4bp"))
temp$bpLocPair <- paste(temp$bpLoc5prime, temp$bpLoc3prime, sep=":")
stem.table <- dcast(temp, coord+region+bpLocPair+type~mutation, value.var = "tileID")
colnames(stem.table)[(ncol(stem.table)-2):ncol(stem.table)] <- c("tileID.mut3p", "tileID.mut5p", "tileID.comp")

temp <- subset(tile.annot, type=="Control")
stem.table$tileID.wt <- temp$tileID[match(stem.table$coord, temp$coord)]

# select domain I and domain II only
stem.table <- subset(stem.table, coord %in% c("hTR:34-190", "hTR:210-366"))

# bp location index
stem.table$bpIdx <- ""
idx <- which(stem.table$type=="Mut1bp")
stem.table$bpIdx[idx] <- gsub(":", ",", stem.table$bpLocPair[idx])
idx <- which(stem.table$type=="Mut4bp")
temp <- strsplit(stem.table$bpLocPair[idx], split=":")
stem.table$bpIdx[idx] <- paste(paste(sapply(temp, "[[", 1), as.numeric(sapply(temp, "[[", 1))+3, sep=":"), paste(as.numeric(sapply(temp, "[[", 2))-3, sapply(temp, "[[", 2), sep=":"), sep=",")

# assign stem ID
stem.table$stemID <- paste0("stem:", 1:nrow(stem.table))
```

Limma for individual replicates
```{r}
stem.rep1 <- runLimma.stemMut(stem.table, "Rep1", 1)
stem.rep2 <- runLimma.stemMut(stem.table, "Rep2", 1)
stem.rep3 <- runLimma.stemMut(stem.table, "Rep3", 1)
```

Merge and p-value based on rank
```{r}
temp <- mergeLimmaResults(list(stem.rep1$limma, stem.rep2$limma, stem.rep3$limma))
stem.table <- cbind(stem.table, temp)
rownames(stem.table) <- rownames(temp)
```

Results
```{r}
# Dotplot of 4bp mutation beta
temp <- subset(stem.table, type=="Mut4bp")
temp$sig <- cut(temp$qval, breaks = c(0,0.1,0.3,1))
table(temp$sig)
temp$effect <- 2^(temp$mean.logFC)
temp$stemID <- paste(temp$region, temp$bpLocPair, sep="_")

p <- ggpubr::ggdotchart(temp, x = "stemID", y = "effect", col="sig", dot.size=3, add = "segments", palette=c("red", "orange", "black"))
p <- p + geom_hline(yintercept = 1, linetype="dashed")
p

# Dotplot of 1bp mutation beta
temp <- subset(stem.table, type=="Mut1bp")
temp$sig <- cut(temp$qval, breaks = c(0,0.1,0.3,1))
table(temp$sig)
temp$effect <- 2^(temp$mean.logFC)
temp$stemID <- paste(temp$region, temp$bpLocPair, sep="_")

p <- ggpubr::ggdotchart(temp, x = "stemID", y = "effect", col="sig", dot.size=2, add = "segments", palette=c("red", "orange", "black"))
p <- p + theme(axis.text.x=element_text(size=6, angle=70, hjust=1, vjust=1))
p <- p + geom_hline(yintercept = 1, linetype="dashed")
p

# Examples: P6a
subset(stem.table, coord=="hTR:210-366" & type=="Mut4bp" & bpLocPair=="257:299")
temp <- stem_log2FC_tabler("stem:108")
p <- ggpubr::ggboxplot(temp, x = "tileID", y = "log2FC", outlier.shape=NA, legend = "none")
p <- p + ggpubr::stat_compare_means(comparisons = list( c("WT", "5' mut"), c("WT", "3' mut"), c("WT", "Comp. mut") ))
p <- p + geom_jitter(aes(col=repl), size=3, width = 0.2, height=0)
p <- p + scale_color_manual(values=c("darkcyan", "black", "brown"))
p <- p + geom_hline(yintercept = 1, linetype="dashed")
p

# Examples: P6.1
subset(stem.table, coord=="hTR:210-366" & type=="Mut4bp" & bpLocPair=="303:313")
temp <- stem_log2FC_tabler("stem:101")
p <- ggpubr::ggboxplot(temp, x = "tileID", y = "log2FC", outlier.shape=NA, legend = "none")
p <- p + ggpubr::stat_compare_means(comparisons = list( c("WT", "5' mut"), c("WT", "3' mut"), c("WT", "Comp. mut") ))
p <- p + geom_jitter(aes(col=repl), size=3, width = 0.2, height=0)
p <- p + scale_color_manual(values=c("darkcyan", "black", "brown"))
p <- p + geom_hline(yintercept = 1, linetype="dashed")
p

# Examples: P5
subset(stem.table, coord=="hTR:210-366" & type=="Mut4bp" & bpLocPair=="243:326")
temp <- stem_log2FC_tabler("stem:92")
p <- ggpubr::ggboxplot(temp, x = "tileID", y = "log2FC", outlier.shape=NA, legend = "none")
p <- p + ggpubr::stat_compare_means(comparisons = list( c("WT", "5' mut"), c("WT", "3' mut"), c("WT", "Comp. mut") ))
p <- p + geom_jitter(aes(col=repl), size=3, width = 0.2, height=0)
p <- p + scale_color_manual(values=c("darkcyan", "black", "brown"))
p <- p + geom_hline(yintercept = 1, linetype="dashed")
p
```

# 5. FlagCTCF-RNA MPRNA-IP

### Load raw count data and tile annotations.
```{r}
load("RawData/git_CTCF_raw.Rdata")
BARCODE_NUM <- 24
REPLICATE_NUM <- 3
```

### Filtering and Pooling
```{r}
pooled.count <- oligoFilter(oligo.count, inputFilterBy="rpm", inputTh=1, ipFilterBy="rpm", ipTh=1) %>% countPooler()
```

```{r}
pooled.fastqc <- fastqc[match(colnames(pooled.count), fastqc$SampleID),]
pooled.fastqc <- droplevels(pooled.fastqc) 
```

### Enrichment

Limma + Voom
```{r}
pooled.model <- model.matrix(data= pooled.fastqc, ~ 0 + Pulldown + Repl)
edgeR.obj <- DGEList(pooled.count)
edgeR.obj <- calcNormFactors(edgeR.obj, method="TMM")
y <- voom(edgeR.obj, pooled.model, plot = T)
fit <- lmFit(y, pooled.model)
contr <- makeContrasts(PulldownIP - PulldownInput, levels = colnames(coef(fit)))
out <- contrasts.fit(fit, contr)
out <- eBayes(out)
out.table <- topTable(out, sort.by = "P", n = Inf)
table(out.table$adj.P.Val<0.01, out.table$logFC>0)
```

Significant tiles
```{r}
tile.table <- data.frame(tileID=tile.annot$tileID, stringsAsFactors = F)
idx <- match(tile.table$tileID, rownames(out.table))
tile.table$rpm <- 2^out.table$AveExpr[idx]
tile.table$log2FC <- out.table$logFC[idx]
tile.table$stat <- out.table$t[idx]
tile.table$pval <- out.table$P.Value[idx]
tile.table$padj <- out.table$adj.P.Val[idx]
tile.table$sig <- factor(tile.table$padj<0.01, levels=c("TRUE", "FALSE"), labels=c("Sig", "Insig"))

tile.table$enriched <- NA
tile.table$enriched[!is.na(tile.table$stat)] <- "Insig"
tile.table$enriched[which(tile.table$sig=="Sig" & tile.table$log2FC>0)] <- "Enriched"
tile.table$enriched[which(tile.table$sig=="Sig" & tile.table$log2FC<0)] <- "Depleted"
table(tile.table$enriched)
```

MA plot
```{r}
p <- ggplot(subset(tile.table, !is.na(log2FC)), aes(x=log10(rpm), y=log2FC))
p <- p + geom_point(data=subset(tile.table, enriched=="Insig"), col="gray", alpha=0.5)
p <- p + geom_point(data=subset(tile.table, enriched=="Depleted"), col="blue", alpha=0.8)
p <- p + geom_point(data=subset(tile.table, enriched=="Enriched"), col="red", alpha=0.8)
p <- p + geom_hline(yintercept = 0, linetype="dashed")
p <- p + xlab("Log10 (Mean RPM)") + ylab("Log2 (IP/Input)")
p
```

### Analysis
par-clip enrichment
```{r}
temp <- tile.table
temp$clip <- tile.annot$clip
subset(tile.annot, is.na(clip))
temp <- subset(temp, !is.na(clip))

temp$clip <- factor(temp$clip, levels=c(0,1), labels=c("No hit", "Hit"))
table(temp$clip)
p <- ggplot(temp, aes(col=clip, log2FC))
p <- p + stat_ecdf(size=1.5)
p
```

2ndary energy predicted by RNAfold
```{r}
temp <- cbind(tile.table, tile.annot[,-1])
temp <- subset(temp, !is.na(stat))
temp$Group <- factor(temp$enriched, levels=c("Depleted", "Insig", "Enriched"))

p <- ggplot(temp, aes(energy, col=Group))
p <- p + stat_ecdf(size=1.5)
p
```

# 6. WRAP53-FlagCTCF MPRNA-IP
### Load raw count data and tile annotations.
```{r}
load("RawData/git_WRAP53_raw.Rdata")
BARCODE_NUM <- 35
REPLICATE_NUM <- 3
```

### Normalization
```{r}
temp <- dcast(oligo.count, tileID+barcodeID~repl+pulldown, value.var="raw")
count.normed <- edgeR::calcNormFactors(edgeR::DGEList(as.matrix(temp[, -c(1,2)])), method="TMM")
count.normed <- edgeR::cpm(count.normed)

temp <- melt(cbind(temp[, c(1,2)], count.normed), id.vars=c("tileID", "barcodeID"), value.name = "normed", variable.name = "sampleID")
idx <- match(paste(oligo.count$tileID, oligo.count$barcodeID, oligo.count$repl, oligo.count$pulldown, sep="_") , paste(temp$tileID, temp$barcodeID, temp$sampleID, sep="_"))
oligo.count$normed <- temp$normed[idx]
```

### Stem mutation

Stem mutant table
```{r}
# construct stem.table using base-pairing index of compensatory mutation
stem.list <- subset(tile.annot, mutation == "compensatory")[ , c("tileID", "mutation", "idx")]
stem.list <- split(stem.list, stem.list$idx)
for (i in 1:length(stem.list)) {
  cur.mut <- subset(tile.annot, idx %in% unlist(strsplit(stem.list[[i]]$idx, split=",")))
  stem.list[[i]] <- rbind(cur.mut[c("tileID", "mutation", "idx")], stem.list[[i]])
}

temp <- do.call(rbind, stem.list)
temp$stemID <- sapply(strsplit(rownames(temp), split="[.]"), "[[", 1)
stem.table <- dcast(temp, stemID~mutation, value.var = "tileID")
colnames(stem.table) <- c("stemID", "tileID.mut3p", "tileID.mut5p", "tileID.comp")

# add wild-types
stem.table$tileID.wt <- subset(tile.annot, type=="wt")$tileID
```

Limma for individual replicates
```{r}
stem.rep1 <- runLimma.stemMut(stem.table, "Rep1", 1)
stem.rep2 <- runLimma.stemMut(stem.table, "Rep2", 1)
stem.rep3 <- runLimma.stemMut(stem.table, "Rep3", 1)
```

Merge and p-value based on rank
```{r}
temp <- mergeLimmaResults(list(stem.rep1$limma, stem.rep2$limma, stem.rep3$limma))
stem.table <- cbind(stem.table, temp)
```

Results
```{r}
temp <- stem.table
temp$sig <- cut(temp$qval, breaks = c(0,0.1,0.3,1))
table(temp$sig)

p <- ggpubr::ggdotchart(temp, x = "stemID", y = "mean.logFC", col="sig", dot.size=4, add = "segments", palette=c("red", "orange", "black"))
p <- p + geom_hline(yintercept = 0, linetype="dashed")
p <- p + ylim(-0.1, 0.2)
p <- p + xlab("Stem location") + ylab("Stem score")
p
```

### Example
```{r}
temp <- stem_log2FC_tabler("56:62,66:72")
p <- ggpubr::ggboxplot(temp, x = "tileID", y = "log2FC", outlier.shape=NA, legend = "none")
p <- p + ggpubr::stat_compare_means(comparisons = list( c("WT", "5' mut"), c("WT", "3' mut"), c("WT", "Comp. mut") ))
p <- p + geom_jitter(aes(col=repl), size=3, width = 0.2, height=0)
p <- p + scale_color_manual(values=c("darkcyan", "black", "brown"))
p <- p + geom_hline(yintercept = mean(subset(temp, tileID=="WT")$log2FC), linetype="dashed")
p
```

### Single mutation
Single mutant table
```{r}
temp <- subset(tile.annot, mutation %in% c("complement", "5p", "3p"))
single.table <- data.frame(mutID=temp$idx, tileID.mut=temp$tileID)
single.table$tileID.wt <- subset(tile.annot, type=="wt")$tileID
```

Limma for individual replicates
```{r}
single.rep1 <- runLimma.singleMut(single.table, "Rep1", 1)
single.rep2 <- runLimma.singleMut(single.table, "Rep2", 1)
single.rep3 <- runLimma.singleMut(single.table, "Rep3", 1)
```

Merge and p-value based on rank
```{r}
temp <- mergeLimmaResults(list(single.rep1$limma, single.rep2$limma, single.rep3$limma))
single.table <- cbind(single.table, temp)
```

Results
```{r}
temp <- single.table
temp$sig <- cut(temp$qval, breaks = c(0,0.1,0.3,1))
table(temp$sig)
temp$region <- sapply(strsplit(temp$tileID.mut, split="_"), "[[", 1)

p <- ggpubr::ggdotchart(temp, x = "mutID", y = "mean.logFC", col="sig", shape="region", dot.size=4, add = "segments", palette=c("red", "orange", "black"))
p <- p + geom_hline(yintercept = 0, linetype="dashed")
p
```


# R session information
```{r}
sessionInfo()
```

